Parallel for Oregon Zip and WRF_Grid Intersection Proportion

### 2017-03-31 r_scripts/parallel_presentation.Rmd
add pic

**Goal:** Estimate population-weighted PM2.5 for each ZIP code.
<br/>
**Problem:** The population data is for WRF Grid, not ZIP code.
<br/>
**Preparation:** 2013 Oregon Zip code Shapefile and WRF Grid
<br/>
**First Step:** Calculate the proportion of intersection of ZIP code and grid box
*The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.*
```{r overlay, echo=FALSE}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

@@

## Parallel on server
**For loop vs Apply**

```{r prepare, include = FALSE}
or_zip_name <- or_zip
length(or_zip_name)
wrf_grid_name <- as.character(smoke_grid@data$WRFGRID_ID)
length(wrf_grid_name)
# empty matrix
zip_wrf_proportion <- matrix(nrow = 417, ncol = 1610, byrow = T,
                             dimnames = list(or_zip_name, wrf_grid_name))

vars1 <- c(1:1610)
z <- rep(201:300, each=100)
w <- rep(201:300,100)

# Setup for parallel computing before for loop ---------------------------------
cores <- detectCores() # 48
cl <- makeCluster(cores) # use half the cores on the vet cluster
registerDoParallel(cl)
# load packages on each cluster
clusterCall(cl, function() library(rgdal))
clusterCall(cl, function() library(sp))
clusterCall(cl, function() library(rgeos))

@@

```
**1. Forloop** (one core, or not parallel computing)

```{r forloop, echo = FALSE}
start <- proc.time()
for (i in 201:300) {
  # output value of zipcode
  zipcode <- or_zip_name[i]
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)
for(j in 201:300){
    # output each grid and create a polygon
    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
    wrf_poly <- SpatialPolygons(wrf_grid@polygons)

@@

```{r mapply, echo = FALSE}
start <- proc.time()
f1 <- function(x){
  zipcode <- as.character(x) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)
}
zip_polygon <- lapply(or_zip_name, f1) # list

@@

Another important thing:
Check if they are identical!
```{r, echo=FALSE, warning = FALSE, include = FALSE}
## write matrix of for loop formed
zip_proportion_df <- data.frame(zip_wrf_proportion)

>>>>

**Goal:** Estimate population-weighted PM2.5 for each ZIP code.
<br/>
**Preparation:** 2013 Oregon Zip code Shapefile and WRF Grid
<br/>
**Problem:** The population data is for WRF Grid, not ZIP code.
<br/>
**First Step:** Calculate the proportion of intersection of ZIP code and grid box
*The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.*
```{r overlay, echo=FALSE}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

@@

## Parallel on server
![serial to solve problems](./serialProblem.jpg)

![Parallel to solve problems](./parallelProblem.jpg)

## Comparisons
**For loop vs Apply**

```{r prepare}
or_zip_name <- or_zip
length(or_zip_name)
wrf_grid_name <- as.character(smoke_grid@data$WRFGRID_ID)
length(wrf_grid_name)
# empty matrix
zip_wrf_proportion <- matrix(nrow = 417, ncol = 1610, byrow = T,
                             dimnames = list(or_zip_name, wrf_grid_name))

vars1 <- c(1:1610)
z <- rep(201:300, each=100)
w <- rep(201:300,100)
```

```{r parallel set}
# Setup for parallel computing before for loop ---------------------------------
cores <- detectCores() # 48
cl <- makeCluster(cores) # use half the cores on the vet cluster
registerDoParallel(cl)
# load packages on each cluster
clusterCall(cl, function() library(rgdal))
clusterCall(cl, function() library(sp))
clusterCall(cl, function() library(rgeos))

@@

**1. Forloop** (one core, or not parallel computing)

```{r forloop}
start <- proc.time()
for (i in 201:300) {
  # output value of zipcode
  zipcode <- or_zip_name[i]
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)
for(j in 201:300){
    # output each grid and create a polygon
    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
    wrf_poly <- SpatialPolygons(wrf_grid@polygons)

@@

```{r mapply}
start <- proc.time()
f1 <- function(x){
  zipcode <- as.character(x) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)
}
zip_polygon <- lapply(or_zip_name, f1) # list

@@

## Another important thing:
Check if they are identical!
```{r, echo=FALSE, warning = FALSE, include = FALSE, message = FALSE}
## write matrix of for loop formed
zip_proportion_df <- data.frame(zip_wrf_proportion)
